---
title: "LandMine Manual"
subtitle: "v.`r SpaDES.core::moduleMetadata(module = 'LandMine', path = '..')$version`"
date: "18 May 2018; updated 6 Sep 2022"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: sandstone
    number_sections: false
    df_print: paged
    keep_md: yes
editor_options:
  chunk_output_type: console
  bibliography: citations/references_LandMine.bib
citation-style: citations/ecology-letters.csl
link-citations: true
always_allow_html: true
---

# *LandMine* Module

<!-- the following are text references used in captions for LaTeX compatibility -->
(ref:LandMine) *LandMine*

```{r setup-LandMine, echo=FALSE, eval=TRUE, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold") ## change to eval = TRUE if all chunks are to be evaluated

## get citation style
if (!file.exists("citations/ecology-letters.csl")) {
  dir.create("citations", showWarnings = FALSE)
  download.file("https://www.zotero.org/styles/ecology-letters", destfile = "citations/ecology-letters.csl")
}

options(repos = c(CRAN = "https://cloud.r-project.org"))

## packages
if (!"pak" %in% rownames(installed.packages())) {
  install.packages("pak")
}

if (!require("dplyr", quietly = TRUE)) {
  pak::pkg_install("dplyr")
}

if (!"SpaDES.core" %in% rownames(installed.packages())) {
  pak::pkg_install( "PredictiveEcology/SpaDES.core@development")
}

if (!"SpaDES.project" %in% rownames(installed.packages())) {
  pak::pkg_install( "PredictiveEcology/SpaDES.project@transition")
}

## package SDM tools unavailable on CRAN since 2020-01-12; use latest archived version
if (!"SDMTools" %in% rownames(installed.packages())) {
  install.packages("https://cloud.r-project.org/src/contrib/Archive/SDMTools/SDMTools_1.1-221.2.tar.gz",
                   repos = NULL)
}

library("Require")
Require(c("dplyr", "igraph", "kableExtra", "raster", "rmarkdown", "SpaDES.core"),
        install = FALSE, update = FALSE)

workingDir <- SpaDES.project::findProjectPath()
moduleDir <- file.path(workingDir, "m") |> checkPath()
inputDir <- file.path(workingDir, "inputs") |> checkPath(create = TRUE)
outputDir <- file.path(workingDir, "outputs") |> checkPath(create = TRUE)
cacheDir <- file.path(outputDir, "cache_LandMineRmd") |> checkPath(create = TRUE)
```

:::{.rmdwarning}
This documentation is work in progress.
Please report any discrepancies or omissions at <https://github.com/PredictiveEcology/LandMine/issues>.
:::

#### Authors:

`r paste(as.character(SpaDES.core::moduleMetadata(module = "LandMine", path = '..')$authors), sep = ', ')`
<!-- ideally separate authors with new lines, '\n' not working -->

## Module Overview

### Module summary

Landmine is a model created for simulating the natural range of variation for landscapes in the boreal forest [@Andison:1996; @Andison:1998].
It has been widely used by the public and the private sector for various purposes.
This SpaDES module is a rewrite of the fire component in native R.

### Model Differences

The current version has not yet been fully tested and compared with the original version, but there are currently several known differences:

1. Fire sizes are taken from a Truncated Pareto distribution, resulting in numerous very small fires, and few large fires;
2. Parameters have been fitted to the landscapes that are under study in the LandWeb project.

### Known Species

Landmine requires the following codes as inputs (the genus and species codes below), which converts and groups species as follows.
Each of the species groups has its own Rate of Spread (ROS) for fire spreading:

```{r species-table, echo=FALSE, eval=TRUE}
spp_df <- data.frame(
  Species = c("Jack pine", "Lodgepole pine", "Unspecified pine species",
              "Paper birch", "Balsam poplar", "Trembling aspen", "Larch/Tamarack",
              "Black spruce",
              "White spruce",
              "Fir species"),
  Group = c("Pine (PINU)", "Pine (PINU)", "Pine (PINU)",
            "Deciduous (DECI)", "Deciduous (DECI)", "Deciduous (DECI)", "Deciduous (DECI)",
            "Black spruce (PICE_MAR)",
            "White spruce (PICE_GLA)",
            "Fir (ABIE)"),
  Code = c("Pinu_ban", "Pinu_con", "Pinu_sp",
           "Betu_pap", "Popu_bal", "Popu_tre", "Lari_lar",
           "Pice_mar",
           "Pice_gla",
           "Abie_sp")
)
knitr::kable(spp_df,
             caption = "LandMine species codes.")
```

\newpage

### Module inputs and parameters

Table \@ref(tab:moduleInputs-LandMine) shows the full list of module inputs.

```{r moduleInputs-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_inputs <- SpaDES.core::moduleInputs("LandMine", "..")
knitr::kable(df_inputs,
             caption = "List of (ref:LandMine) input objects and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

Summary of user-visible parameters (Table \@ref(tab:moduleParams-LandMine)).

```{r moduleParams-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_params <- SpaDES.core::moduleParams("LandMine", "..")
knitr::kable(df_params, caption = "List of (ref:LandMine) parameters and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

### Module outputs

Description of the module outputs (Table \@ref(tab:moduleOutputs-LandMine)).

```{r moduleOutputs-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_outputs <- SpaDES.core::moduleOutputs("LandMine", "..")
knitr::kable(df_outputs, caption = "List of (ref:LandMine) outputs and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = TRUE)
```

\newpage

## Usage

To run this Landmine module alone (*i.e.*, for fitting), the following should work (*iff* raster inputs for `rstStudyRegion` and `rstFlammable` are available), assuming all R packages are available.

**NB:** Paths will have be changed for a different user.

### Package dependencies

To determine which packages are used by `LandMine`, use:

```{r module-package-deps, eval=TRUE}
SpaDES.core::packages(modules = "LandMine", paths = moduleDir)[[1]]
```

### Module usage

```{r study-area}
studyArea <- SpaDES.tools::randomStudyArea(seed = 1234, size = 1e10)
rasterToMatch <- raster(studyArea, res = 250)
```

```{r sim-setup}
times <- list(start = 0, end = 13)

parameters <- list(
  # LandMine = list(flushCachedRandomFRI = TRUE) 
)

modules <- list("LandMine")

objects <- list(
  studyArea = studyArea,
  rasterToMatch = rasterToMatch
)

paths <- list(
  cachePath = cacheDir,
  modulePath = moduleDir,
  inputPath = inputDir,
  outputPath = outputDir
)
```

```{r run-sim}
mySim <- simInit(times = times, params = parameters, modules = modules,
                 objects = objects, paths = paths)

dev()
mySimOut <- spades(mySim, .plotInitialTime = times$start, debug = TRUE)
```

\newpage

## Testing the burn algorithm

```{r testing}
Require(c("data.table", "DEoptim", "parallel", "SDMTools"))

s <- simInit(times = times, params = parameters, modules = modules,
             objects = objects, paths = paths)

ros <- raster(extent(0, 2.5e5, 0, 2.5e5), res = 250, vals = 0)
ros <- ros == 0
fireSize <- 1e5
```

```{r functions}
source(file.path(moduleDir, "LandMine", "R", "burn.R"))

burnFun <- function(ros, centreCell, fireSize, spawnNewActive, sizeCutoffs, burn1, spreadProb) {
    burned <- burn1(
      landscape = ros,
      startCells = centreCell,
      fireSizes = fireSize,
      spreadProb = spreadProb,
      spreadProbRel = ros,
      spawnNewActive = spawnNewActive,
      sizeCutoffs = sizeCutoffs
    )
    burnedMap <- raster(ros)
    burnedMap[] <- NA
  
    burnedMap[burned$pixels] <- burned$initialPixels
    LM <- SDMTools::PatchStat(burnedMap, cellsize = res(burnedMap)[1])
    list(burnedMap = burnedMap, LM = LM)
}

makeParallel <- function(wantParallel, numClus, cl, funs, addlPkgs) {
  if (wantParallel) {
    library(parallel)
    if (is.null(cl)) {
      cl <- parallel::makeCluster(numClus)
    }
    funs <- c(funs)
    parallel::clusterExport(cl,  funs)
    #env <- environment()
    parallel::clusterExport(cl, "addlPkgs", envir = parent.frame())
    parallel::clusterEvalQ(cl, {lapply(addlPkgs, library, character.only = TRUE)})
    return(cl)
  }
}

wrap <- function(cl = NULL, reps, par, burn1, burnFun, objs, addlPkgs) {
  funNames <- c(deparse(substitute(burn1)), deparse(substitute(burnFun)))
  cl <- makeParallel(TRUE,
                     numClus = if (length(cl)) length(cl) else detectCores() - 1,
                     cl, c(funNames, objs), addlPkgs)
  objs <- append(mget(objs, envir = parent.frame()), list(burn1 = burn1))
  parallel::clusterExport(cl,  "objs", envir = parent.frame())
  burnMapList <- parallel::clusterApplyLB(cl, reps, function(r) {
    do.call("burnFun", objs)
  })
  return(list(cl = cl, out = burnMapList))
}

fitSN <- function(sna, ros, centreCell, fireSizes = 10^(2:5),
                  desiredPerimeterArea = 0.004, burnFun, burn1) {
  sizeCutoffs <- 10^sna[5:6]
  spreadProb <- sna[7]
  sna <- c(10^(sna[1]), 10^(sna[2]), 10^(sna[3]), 10^(sna[4]))
  bfs1 <- lapply(fireSizes, function(fireSize) {
    burnFun(ros, centreCell, fireSize, 
            sna, sizeCutoffs, burn1 = burn1, spreadProb = spreadProb)
  })
  res <- lapply(seq(bfs1), function(bfCount) {
    abs(log(bfs1[[bfCount]]$LM[1,"perim.area.ratio"]) -
          log(desiredPerimeterArea)) + 
      100*(sum(bfs1[[bfCount]]$burnedMap[], na.rm = TRUE) < fireSizes[bfCount])
    # it needs to get to above 90,000 HA for it to count
  })
  a <- sum(unlist(res))# * log10(fireSizes)) # weigh larger ones more
  attr(a, "bfs1") <- bfs1
  a
}

fitSN2 <- function(par, ros, centreCell, fireSizes = 10^(2:5), 
                   desiredPerimeterArea = 0.003, spreadProb = 0.9) {
  sizeCutoffs <- 10^c(par[4], par[5])
  bfs1 <- lapply(fireSizes, function(fireSize) {
    sna <- min(-0.15, par[1] + par[2]*log10(fireSize))
    sna <- 10^c(sna*par[3], sna*2*par[3], sna*3*par[3], sna*4*par[3])
    #sna <- -1
    burnFun(ros, centreCell, fireSize, sna, sizeCutoffs, burn1, spreadProb)
  })
  res <- lapply(seq(bfs1), function(bfCount) {
    abs(log(bfs1[[bfCount]]$LM[1, "perim.area.ratio"]) -
          log(desiredPerimeterArea)) + 
      100*(sum(bfs1[[bfCount]]$burnedMap[], na.rm = TRUE) < fireSizes[bfCount])
    # it needs to get to above 90,000 HA for it to count
  })
  a <- sum(unlist(res))
  attr(a, "bfs1") <- bfs1
  a
}
```

\newpage

## Optimizing parameters

The following code chunk tries to find values of `spawnNewActive` that creates "reasonable" fire shapes at all sizes.

```{r fit-values}
wantParallel <- TRUE
maxRetriesPerID <- 4 ## 4 retries (5 attempts total)
spreadProb <- 0.9
spawnNewActive <- c(0.46, 0.2, 0.26, 0.11)
sizeCutoffs <- c(8e3, 2e4)
#spawnNewActive <- c(0.1, 0.04, 0.025, 0.01)
#sizeCutoffs <- c(8e3, 2e4)

NineCorners <- cellFromRowCol(ros,
                              row = nrow(ros) / 4 * rep(1:3, 3),
                              col = ncol(ros) / 4 * rep(1:3, each = 3))

centreCell <- cellFromRowCol(ros,
                             row = nrow(ros) / 2,
                             col = ncol(ros) / 2)

## Set variables
objs <- c("ros", "centreCell", "fireSize",
          "spawnNewActive", "sizeCutoffs", "spreadProb")
funs <- c("burnFun", "burn1")
addlPkgs <- c("data.table", "raster", "SDMTools", "SpaDES.tools")
```

```{r make-cluster}
### SET UP CLUSTER FOR PARALLEL
numCores <- min(detectCores() / 2, 120L)
if (Sys.info()["sysname"] == "Windows") {
  cl <- parallel::makeCluster(numCores)
} else if (grepl("W-VIC-A1053|pinus[.]for-cast[.]ca", Sys.info()["nodename"])) {
  Require("future")
  
  ## TODO: customize this to make a cluster on several machines via ssh
  clNames <- c(
    rep("localhost", 0),
    rep("picea.for-cast.ca", 25),
    rep("pseudotsuga.for-cast.ca", 80)
  )
  
  ## check the cluster:
  ## 1. at least 10 populations per parameter; 7 params;
  ## 2. ensure it's a multiple of the number of params (7);
  ## 3. R can't create more than ~125 socket connections.
  stopifnot(
    length(clNames) >= 70,
    length(clNames) %% 7 == 0,
    length(clNames) <= 120
  )
  cl <- makeClusterPSOCK(clNames, homogeneous = FALSE, verbose = TRUE)
  
  clusterSetRNGStream(cl, sample(1e8, 1))
} else {
  cl <- parallel::makeCluster(numCores, type = "FORK")
  clusterSetRNGStream(cl, sample(1e8, 1))
}

if (!inherits(cl[[1]], "forknode")) {
  parallel::clusterExport(cl, funs)
  parallel::clusterExport(cl, objs )
  #env <- environment()
  parallel::clusterExport(cl,  "addlPkgs", envir = parent.frame())
  parallel::clusterEvalQ(cl, {
    lapply(addlPkgs, library, character.only = TRUE)
  })
}
```

```{r optimization}
opt_sn <- DEoptim(fitSN,
                  lower = c(-2, -3, -3, -3, 1, 3.5, 0.75), 
                  upper = c(-0.1, -0.5, -0.5, -1, 3.5, 5, 1),
                  control = DEoptim.control(VTR = 0.001, NP = as.integer(length(clNames)),
                                            itermax = 200, cluster = cl, strategy = 6),
                  ros = ros,
                  centreCell = centreCell,
                  fireSizes = c(10, 100, 1000, 10000, 100000),
                  desiredPerimeterArea = 0.003,
                  burnFun = burnFun,
                  burn1 = burn1)

opt_sn$optim$bestmem ## best param values
# dput(unname(opt_sn$optim$bestmem))

## assign with suffix to facilitate multiple DEoptim runs
timestamp <- format(Sys.time(), "%Y%M%d%H%M%S")
assign(paste0("opt_sn_", timestamp), opt_sn)
saveRDS(opt_sn, file.path(moduleDir, "LandMine", "data", paste0(timestamp, "_DEoptim_250m.rds")))
```

```{r fitSn}
fs_sn <- c(10, 100, 1000, 10000, 100000)
fit_sn <- fitSN(sna = c(-1, -1, -1, -2, 2, 4, 0.9),
                ros = ros,
                centreCell = centreCell,
                fireSizes = fs_sn,
                desiredPerimeterArea = 0.003,
                burnFun = burnFun,
                burn1 = burn1)
bfs1_sn <- purrr::transpose(attr(fit_sn, "bfs1"))
LM_sn <- do.call(rbind, bfs1_sn$LM)
plot(fs_sn, LM_sn[, "perim.area.ratio"]) ## NOTE: visual inspection - not too round; not too sinuous
```

A second (alternative) version tries the optimization using fewer parameters, to test whether a simpler version gets better/different results.
Although this version was not used for the final module, we preserve it here for posterity.

```{r optimization2}
fs_optim2 <- c(0.2, 1:8)*10000
opt_sn2 <- DEoptim(fitSN2, lower = c(1, -1, 1, 3, 4), upper = c(3, -0.3, 3, 4, 5), 
                   control = DEoptim.control(VTR = 0.001, itermax = 40, 
                                             cluster = cl, strategy = 6), 
                   ros = ros, centreCell = centreCell, 
                   fireSizes = fs_optim2, desiredPerimeterArea = 0.003)
```

```{r relationship-btwn-perimArea-and-fireSize}
fs_sn2 <- round(runif(10, 10, 4000))
fit_sn2 <- fitSN2(par = c(2, -0.63333, 1, 3.2, 4.4),
                  ros = ros,
                  centreCell = centreCell,
                  fireSizes = fs_sn2,
                  desiredPerimeterArea = 0.003,
                  spreadProb = 0.9)
bfs1_sn2 <- purrr::transpose(attr(fit_sn2, "bfs1"))
LM_sn2 <- do.call(rbind, bfs1_sn2$LM)
plot(fs_sn2, LM_sn2[, "perim.area.ratio"]) ## NOTE: visual inspection - not too round; not too sinuous
```

\newpage

## Manual inspection of optimization results

### Original (2018) version

The original version was run using 100m pixels, despite the simulations being run using 250m pixels.
This was corrected and rerun below.

```{r manual-run-2018}
## 10,000 hectares burns gave this
spawnNewActive[2:3] <- c(0.0235999945606232, 0.0263073265505955)

#100,000 hectare burns gave this
#spawnNewActive <- 10^c(-1.264780,   -1.970946,   -1.288213,   -2.202580)
spawnNewActive <- 10^c(-0.586503645288758, -1.08108837273903,
                       -2.14391896536108, -1.00221184641123)
sizeCutoffs <- 10^c(3.37711253212765, 4.52040993282571)

sns <- c(-1.733262, -0.933318, -2.562183, -2.493687, 3.064458, 4.812305)
spawnNewActive <- 10^sns[1:4]
sizeCutoffs <- 10^sns[5:6]
#spawnNewActive <- 10^c(-1.646419, -1.815395, -2.809013, -2.613337)
#sizeCutoffs <- 10^c(3.888317,    4.641961)

## 100,000 pixel fires -- the next worked, but I think we can get better
# sns <- structure(
#   c(-1.652459,-0.962121,-0.964879,-2.304902, 3.522345, 4.173242),
#   .Names = c("par1", 
sns <- structure(
  c(-1.641197,-1.152821,-0.697335,-1.751917, 3.720378, 4.034059),
  .Names = c("par1", "par2", "par3", "par4", "par5", "par6")
)
spawnNewActive <- 10^sns[1:4]
sizeCutoffs <- 10^(sns[5:6])
fireSize <- 30000

## 100
sns <- c(-0.77716149196811, -0.769325340166688, -1.2772046867758,
         -1.99332102853805, 3.14260408212431, 4.46155184064992) 

## 1000
sns <- c(-0.775107,-1.031760,-0.599669,-1.958105, 3.048958, 4.275831)

## seemed good for 100,000, pretty good for 1e3
sns <- c(-1.54885, -0.97052, -1.38305, -1.93759, 3.20379, 4.13237)

## good for 100 000, 10 000 ha -- too sinuous for 1000 and 100 ha
sns <- c(-1.537203,-1.462981,-0.524957,-1.002567, 3.642046, 4.501754)

## good for 100 000, 10 000 ha (except some fires @ 1e5 don't make it to full size)
## -- too sinuous for smaller
sns <- c(-1.484338,-1.220440,-2.948275,-2.155940, 3.945281, 4.904893)

sns <- c(-1.495247,-0.800494,-1.582350,-2.270646, 3.530671, 4.663245)

## final optimization after 75 iterations, Good: 1e5, 1e4
sns <- c(-1.47809, -0.86224, -1.34532, -1.93568, 3.27149, 4.20741)

## based on equal weights 10^(1:5)
sns <- c(-0.923528, -1.804549, -1.760455, -1.793594,  1.683355,  4.466668)

## With spreadProb = 0.9 # Pretty GOOD!
sns <- c(-0.731520, -0.501823, -0.605968, -1.809726,  2.202732,  4.696060, 0.9) ## used in module

## With spreadProb = 0.9 # Optimal
sns <- c(-0.978947, -0.540946, -0.790736, -1.583039,  2.532013,  4.267547,  0.946730)

spawnNewActive <- 10^sns[1:4]
sizeCutoffs <- 10^(sns[5:6])
if (length(sns) == 7) spreadProb <- sns[7]

# from linear model version
par <- c(1.548899,-0.396904, 2.191424, 3.903082, 4.854002)
sizeCutoffs <- 10^c(par[4], par[5])
sna <- min(-0.15, par[1] + par[2]*log10(fireSize))
sna <- 10^c(sna*par[3], sna*2*par[3], sna*3*par[3], sna*4*par[3])
spawnNewActive <- sna

###########################
clearPlot()
dev()

for (i in 1:5) {
  fireSize <- 10^i
  dim <- round(sqrt(fireSize)*5 * 250)
  ros <- raster(extent(0, dim,0,dim), res = 250, vals = 1)
  centreCell <- cellFromRowCol(ros,
                               rownr = nrow(ros) / 2,
                               colnr = ncol(ros) / 2)
  reps <- paste0("rep", 1:4 + (log10(fireSize) - 1)*4)
  burnedMapList <- wrap(cl = cl, reps = reps, par = TRUE, burn1 = burn1,
                        burnFun = burnFun, objs = objs, libs = addlPkgs)
  names(burnedMapList$out) <- reps
  burnedMapList <- purrr::transpose(burnedMapList$out)
  do.call(rbind, burnedMapList$LM)
  
  Plot(burnedMapList$burnedMap, cols = c("red"), new = FALSE, na.color = "white",
       legend = FALSE, visualSqueeze = 0.7, title = paste0("Fire size: ", fireSize))
}
```

```{r another_option}
reps <- paste0("rep", 1:1)
perims <- list()
perm <- list()
mod <- list()
dev()
clearPlot()
fireSizes <-  10^(4)

for (fs in fireSizes) {
  for (i in 1:1) {
    ros <- raster(extent(0, 2e5, 0, 2e5), res = 250, vals = 1)
    NineCorners <- cellFromRowCol(
      ros,
      rownr = nrow(ros) / 4 * rep(1:3, 3),
      colnr = ncol(ros) / 4 * rep(1:3, each = 3)
    )
    centreCell <- NineCorners
    ran <- runif(4, -3, -1)
    spawnNewActive <- 10^ran
    #spawnNewActive <- 10^c(-0.1, -0.75, -1.2, ran*2.5)
    fireSize = rep(fs, length(centreCell))
    sizeCutoffs <- 10^c(1,3)
    burnedMapList <- wrap(cl = cl, reps = reps, par = TRUE, burn1 = burn1,
                          burnFun = burnFun, objs = objs, libs = addlPkgs)
    names(burnedMapList$out) <- reps
    burnedMapList <- purrr::transpose(burnedMapList$out)
    do.call(rbind, burnedMapList$LM)
    Plot(burnedMapList$burnedMap, new = TRUE, zero.color = "white")
    perims[[i]] <- data.frame(perim = burnedMapList$LM$rep1$perim.area.ratio,
                              spawnNewActive = mean(spawnNewActive),
                              others = t(spawnNewActive))
  }
  fsChar <- as.character(fs)
  perm[[fsChar]] <- rbindlist(perims)
  perm[[fsChar]]$perim <- log10(perm[[fsChar]]$perim)
  perm[[fsChar]]$spawnNewActive <- perm[[fsChar]]$spawnNewActive
  Plot(perm[[fsChar]]$perim, perm[[fsChar]]$spawnNewActive, new = TRUE,
       addTo = paste0("fs",fsChar))
  mod[[fsChar]] <- lm(spawnNewActive ~ perim, data = perm[[fsChar]])
}

(predict(mod[["10"]], data.frame(perim = log10(0.003))))
(predict(mod[["100"]], data.frame(perim = log10(0.003))))
log10(predict(mod[["1000"]], data.frame(perim = log10(0.003))))
```

### Current (2022) version

The original version was run using 100m pixels, despite the simulations being run using 250m pixels.
This version uses 250m pixels.

```{r manual-run-2022}
## final optimization after 170 iterations (fitSN)
sns <- unname(opt_sn$optim$bestmem)

spawnNewActive <- 10^sns[1:4]
sizeCutoffs <- 10^(sns[5:6])
spreadProb <- sns[7]

## from linear model version
par <- c(1.548899,-0.396904, 2.191424, 3.903082, 4.854002)
sizeCutoffs <- 10^c(par[4], par[5])
sna <- min(-0.15, par[1] + par[2]*log10(fireSize))
sna <- 10^c(sna*par[3], sna*2*par[3], sna*3*par[3], sna*4*par[3])
spawnNewActive <- sna

clearPlot()

for (i in 1:5) {
  fireSize <- 10^i
  dim <- round(sqrt(fireSize)*5 * 250)
  ros <- raster(extent(0, dim, 0, dim), res = 250, vals = 1)
  centreCell <- cellFromRowCol(ros,
                               row = nrow(ros) / 2,
                               col = ncol(ros) / 2)
  reps <- paste0("rep", 1:4 + (log10(fireSize) - 1)*4)
  burnedMapList <- wrap(cl = cl, reps = reps, par = TRUE, burn1 = burn1,
                        burnFun = burnFun, objs = objs, addlPkgs = addlPkgs)
  names(burnedMapList$out) <- reps
  burnedMapList <- purrr::transpose(burnedMapList$out)
  do.call(rbind, burnedMapList$LM)
  
  Plot(burnedMapList$burnedMap, cols = c("red"), new = FALSE, na.color = "white",
       legend = FALSE, visualSqueeze = 0.7, title = paste0("Fire size: ", fireSize))
}
```

### Cleaning up

```{r cleanup}
parallel::stopCluster(cl)
unlink(cacheDir, recursive = TRUE)
```

### Code and data availability

Code available from <https://github.com/PredictiveEcology/LandMine>.

### Links to other modules

Originally developed as part of the [LandWeb](https://github.com/PredictiveEcology/LandWeb) project.
