---
title: "LandMine Manual"
subtitle: "v.`r SpaDES.core::moduleVersion(module = 'LandMine', path = '..')`"
date: "18 May 2018; updated 29 Oct 2025"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: sandstone
    number_sections: false
    df_print: paged
    keep_md: yes
editor_options:
  chunk_output_type: console
  bibliography: citations/references_LandMine.bib
citation-style: citations/ecology-letters.csl
link-citations: true
always_allow_html: true
---

# *LandMine* Module

<!-- the following are text references used in captions for LaTeX compatibility -->
(ref:LandMine) *LandMine*

```{r setup-LandMine, echo = FALSE, eval = TRUE, include = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold")

## get citation style
if (!file.exists("citations/ecology-letters.csl")) {
  dir.create("citations", showWarnings = FALSE)
  download.file("https://www.zotero.org/styles/ecology-letters", destfile = "citations/ecology-letters.csl")
}

options(repos = c(CRAN = "https://cloud.r-project.org"))

## packages
if (!"pak" %in% rownames(installed.packages())) {
  install.packages("pak")
}

if (!require("dplyr", quietly = TRUE)) {
  pak::pkg_install("dplyr")
}

if (!"SpaDES.core" %in% rownames(installed.packages())) {
  pak::pkg_install( "PredictiveEcology/SpaDES.core@development")
}

if (!"SpaDES.project" %in% rownames(installed.packages())) {
  pak::pkg_install( "PredictiveEcology/SpaDES.project@development")
}

if (!"LandWebUtils" %in% rownames(installed.packages())) {
  pak::pkg_install("PredictiveEcology/LandWebUtils@development")
}

Require::Require(
  packages = c(
    "dplyr", "DEoptim", "ggplot2", "igraph", "kableExtra",
    "LandWebUtils", "purrr", "rmarkdown", "SpaDES.core", "terra", "tidyterra"
  ),
  install = FALSE,
  update = FALSE
)

if (interactive()) {
  workingDir <- SpaDES.project::findProjectPath() ## e.g., ~/GitHub/LandWeb
  moduleDir <- file.path(workingDir, "modules") |> fs::path_norm()
} else {
  ## e.g., when rendering module Rmd
  workingDir <- file.path("..") |> fs::path_norm() |> dirname()
  moduleDir <- file.path("..") |> fs::path_norm()
}
inputDir <- file.path(workingDir, "inputs") |> fs::dir_create()
outputDir <- file.path(workingDir, "outputs") |> fs::dir_create()
cacheDir <- file.path(outputDir, "cache_LandMineRmd") |> fs::dir_create()
figDir <- file.path(moduleDir, "LandMine", "figures") |>
    fs::dir_create()
```

:::{.rmdwarning}
This documentation is work in progress.
Please report any discrepancies or omissions at <https://github.com/PredictiveEcology/LandMine/issues>.
:::

#### Authors:

`r paste(as.character(SpaDES.core::moduleMetadata(module = "LandMine", path = '..')$authors), sep = ', ')`
<!-- ideally separate authors with new lines, '\n' not working -->

## Module Overview

### Module summary

Landmine is a model created for simulating the natural range of variation for landscapes in the boreal forest [@Andison:1996; @Andison:1998].
It has been widely used by the public and the private sector for various purposes.
This SpaDES module is a rewrite of the fire component in native R.

### Model Differences

The current version has not yet been fully tested and compared with the original version, but there are currently several known differences:

1. Fire sizes are taken from a Truncated Pareto distribution, resulting in numerous very small fires, and few large fires;
2. Parameters have been fitted to the landscapes that are under study in the LandWeb project.

### Known Species

Landmine requires the following codes as inputs (the genus and species codes below), which converts and groups species as follows.
Each of the species groups has its own Rate of Spread (ROS) for fire spreading:

```{r species-table, echo = FALSE, eval = TRUE}
data.frame(
  Species = c(
    "Jack pine", "Lodgepole pine", "Unspecified pine species",
    "Paper birch", "Balsam poplar", "Trembling aspen", "Larch/Tamarack",
    "Black spruce",
    "White spruce",
    "Fir and other softwoods"
  ),
  Code = c(
    "Pinu_ban", "Pinu_con", "Pinu_spp",
    "Betu_pap", "Popu_bal", "Popu_tre", "Lari_lar",
    "Pice_mar",
    "Pice_gla",
    "Abie_spp"
  ),
  Group = c(
    "Pine (PINU)", "Pine (PINU)", "Pine (PINU)",
    "Deciduous (DECI)", "Deciduous (DECI)", "Deciduous (DECI)", "Deciduous (DECI)",
    "Black spruce (PICE_MAR)",
    "White spruce (PICE_GLA)",
    "Fir (ABIE)"
  )
) |>
  knitr::kable(caption = "LandMine species codes.")
```

\newpage

### Module inputs and parameters

Table \@ref(tab:moduleInputs-LandMine) shows the full list of module inputs.

```{r moduleInputs-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_inputs <- SpaDES.core::moduleInputs("LandMine", "..")
knitr::kable(df_inputs,
             caption = "List of (ref:LandMine) input objects and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = FALSE)
```

Summary of user-visible parameters (Table \@ref(tab:moduleParams-LandMine)).

```{r moduleParams-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_params <- SpaDES.core::moduleParams("LandMine", "..")
knitr::kable(df_params, caption = "List of (ref:LandMine) parameters and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = FALSE)
```

### Module outputs

Description of the module outputs (Table \@ref(tab:moduleOutputs-LandMine)).

```{r moduleOutputs-LandMine, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_outputs <- SpaDES.core::moduleOutputs("LandMine", "..")
knitr::kable(df_outputs, caption = "List of (ref:LandMine) outputs and their description.") |>
  kableExtra::kable_styling(latex_options = "scale_down", full_width = FALSE)
```

\newpage

## Usage

To run this Landmine module alone (*i.e.*, for fitting), the following should work (*iff* raster inputs for `studyArea` and `rasterToMatch` are available), assuming all R packages are available.

**NB:** Paths will have be changed for a different user.

### Package dependencies

To determine which packages are used by `LandMine`, use:

```{r module-package-deps, eval = TRUE}
SpaDES.core::packages(modules = "LandMine", paths = "..")[[1]]
```

### Module usage

First, define a study area and create a template raster.

```{r study-area}
studyArea <- SpaDES.tools::randomStudyArea(seed = 1234, size = 1e10)
rasterToMatch <- terra::rast(studyArea, resolution = 240)
```

Next, set up the simulation to run for 13 timesteps using default module parameters.

```{r sim-setup}
times <- list(start = 0, end = 13)

parameters <- list()

modules <- list("LandMine")

objects <- list(
  studyArea = studyArea,
  rasterToMatch = rasterToMatch
)

paths <- list(
  cachePath = cacheDir,
  modulePath = moduleDir,
  inputPath = inputDir,
  outputPath = outputDir
)
```

```{r run-sim}
mySim <- simInit(
  times = times,
  params = parameters,
  modules = modules,
  objects = objects,
  paths = paths
)

dev()
mySimOut <- spades(mySim, .plotInitialTime = times$start, debug = TRUE)
```

\newpage

## Testing the burn algorithm

```{r testing}
s <- simInit(
  times = times, 
  params = parameters, 
  modules = modules,
  objects = objects, 
  paths = paths
)
```

\newpage

## Optimizing parameters

The following code chunk tries to find values of `spawnNewActive` that creates "reasonable" fire shapes at all sizes.

```{r fit-values}
pixelSize <- 240
DEoptim_file <- file.path(moduleDir, "LandMine", "data", "LandMine_DEoptim_params.csv")

## The calibration landscape is deliberately uniform and fully flammable, so the fitted
## shape statistics aren't confounded by any one study area's fuel pattern. The helper
## also returns the ignition cell and a file path -- the objective takes the landscape as
## a path because `SpatRaster`s can't be serialized to cluster workers.
lscape <- LandWebUtils::landmine_optim_landscape(pixelSize = pixelSize)
ros <- lscape$file
centreCell <- lscape$centreCell

## Fire sizes are specified in HECTARES, because that is how the calibration targets are
## defined (Andison fitted the shape relationship over 10-3,000 ha), and converted to pixels
## for the resolution being calibrated. Specifying pixel counts directly means a different
## physical fire at every resolution -- `c(10, ..., 1e5)` pixels is 10-100,000 ha at 100 m but
## 57.6-576,000 ha at 240 m -- which put most of the v3 calibration outside the fitted range,
## and defeats the point of calibrating at several pixel sizes.
fireSizesHa <- c(50, 200, 600, 1500, 3000)
fireSizes <- LandWebUtils::landmine_optim_fireSizes(pixelSize, fireSizesHa)

## Current best-fit parameters, used by the visual-inspection runs further below.
## Read via `landmine_optim_params_read()`/`landmine_optim_unpack()` so the CSV schema and
## the `10^` convention have a single definition, shared with `LandMine.R`.
nRuns <- nrow(LandWebUtils::landmine_optim_params_read(DEoptim_file))
pars <- LandWebUtils::landmine_optim_unpack(
  LandWebUtils::landmine_optim_params_read(DEoptim_file, rowID = nRuns)
)
spawnNewActive <- pars$spawnNewActive
sizeCutoffs <- pars$sizeCutoffs
spreadProb <- pars$spreadProb
fireSize <- 1e5

## Set variables to be exported to the cluster workers
objs <- c("centreCell", "fireSize", "ros", "spawnNewActive", "sizeCutoffs", "spreadProb")
pkgs <- c("data.table", "LandWebUtils", "terra", "SpaDES.tools")
```

```{r make-cluster}
wantParallel <- TRUE

## numCores should be >= 70 and needs to be multiple of number of params to be fit (7)
numCores <- (parallelly::availableCores(constraints = c("connections")) %/% 7) * 7

clNames <- rep("localhost", numCores)

if (wantParallel) {
  cl <- LandWebUtils::landmine_optim_clusterSetup(nodes = clNames)
} else {
  cl <- NULL
}

if (!inherits(cl[[1]], "forknode")) {
  LandWebUtils::landmine_optim_clusterExport(cl, objs = objs, pkgs = pkgs)
}

## double check that necessary objects exist:
if (!is.null(cl)) {
  parallel::clusterEvalQ(cl, ls())
}
```

### Visual examination

```{r optimization}
## `landmine_optim_calibrate()` wraps the DEoptim fit; its defaults reproduce the fit that
## was previously written out inline here (same bounds, VTR, itermax, strategy, objective).
## `objective = "andison"` (the default) fits the shape-vs-area relationship and the remnant
## island fractions from Andison (1996), rather than a single constant perimeter:area target.
## `crnSeed` applies common random numbers, so parameter vectors are compared against the same
## random draws; `replicates` averages over draws, which the island statistic needs.
opt_sn <- LandWebUtils::landmine_optim_calibrate(
  pixelSize = pixelSize,
  fireSizesHa = fireSizesHa,
  itermax = 200,
  cl = cl,
  seed = 1234L,   ## with a fixed NP, this makes the per-worker RNG streams reproducible
  crnSeed = 1L,
  paramsFile = DEoptim_file ## appends the best parameters as a new dated row
)

opt_sn$optim$bestmem ## best param values

parallel::stopCluster(cl)
cl <- NULL
gc()
```

```{r fitSn}
fs_sn <- c(10, 100, 1000, 10000, 100000)
fit_sn <- LandWebUtils::landmine_optim_fitSN(
  sna = c(-1, -1, -1, -2, 2, 4, 0.9),
  ros = ros,
  centreCell = centreCell,
  fireSizes = fs_sn,
  desiredPerimeterArea = 0.003
)
bfs1_sn <- purrr::transpose(attr(fit_sn, "bfs1"))
LM_sn <- do.call(rbind, bfs1_sn$LM)

## NOTE: visual inspection - not too round; not too sinuous
plot(log10(fs_sn), LM_sn[, "perim.area.ratio"])
```

## Alternate optimization

A second (alternative) version tries the optimization using fewer parameters, to test whether a simpler version gets better/different results.
Although this version was not used for the final module, we preserve it here for posterity.

```{r optimization2}
## NOTE: needs a live cluster; re-create it (the `optimization` chunk stops its own).
cl <- LandWebUtils::landmine_optim_clusterSetup(nodes = clNames)

fs_optim2 <- c(0.2, 1:8) * 10000
opt_sn2 <- DEoptim::DEoptim(
  fn = LandWebUtils::landmine_optim_fitSN2,
  lower = c(1, -1, 1, 3, 4),
  upper = c(3, -0.3, 3, 4, 5), 
  control = DEoptim::DEoptim.control(
    VTR = 0.001, 
    itermax = 40, 
    cluster = cl,
    strategy = 6
  ), 
  ros = ros, 
  centreCell = centreCell, 
  fireSizes = fs_optim2, 
  desiredPerimeterArea = 0.003
)
```

### Visual examination

```{r relationship-btwn-perimArea-and-fireSize}
fs_sn2 <- round(runif(10, 10, 4000))
fit_sn2 <- LandWebUtils::landmine_optim_fitSN2(
  par = c(2, -0.63333, 1, 3.2, 4.4),
  ros = ros,
  centreCell = centreCell,
  fireSizes = fs_sn2,
  desiredPerimeterArea = 0.003,
  spreadProb = 0.9
)
bfs1_sn2 <- purrr::transpose(attr(fit_sn2, "bfs1"))
LM_sn2 <- do.call(rbind, bfs1_sn2$LM)

## NOTE: visual inspection - not too round; not too sinuous
plot(log10(fs_sn2), LM_sn2[, "perim.area.ratio"]) 
```

\newpage

## Manual inspection of optimization results

Fires are simulated at a range of sizes and replicated, then plotted for visual
inspection: the shapes should be neither too round nor too sinuous. This uses 240 m pixels,
matching LandWeb v3. (Earlier 100 m and 250 m explorations are in the git history.)

```{r manual-run-2025}
library(ggplot2)
library(tidyterra)

## `spawnNewActive`, `sizeCutoffs` and `spreadProb` are set in the `fit-values` chunk
## above, from the most recent row of the calibration CSV.

cl <- NULL ## new cluster created below; will be reused for each fireSize
n <- 3^2 ## number of reps per fireSize (square numbers best for plotting)

for (i in 1:5) {
  fireSize <- 10^i
  ## landscape sized to the fire (5x the fire's linear extent), via the shared helper
  lscape_i <- LandWebUtils::landmine_optim_landscape(
    pixelSize = pixelSize,
    n = round(sqrt(fireSize) * 5)
  )
  ros <- lscape_i$file
  centreCell <- lscape_i$centreCell
  reps <- sprintf("rep%02d", 1:n + (i - 1)*n)
  burnedMapList <- LandWebUtils::landmine_optim_clusterWrap(
    cl = cl,
    nodes = clNames, 
    reps = reps, 
    objs = objs, 
    pkgs = pkgs
  )
  cl <- burnedMapList$cl ## recover the cluster for re-use in the loop
  
  names(burnedMapList$out) <- reps
  burnedMapList <- purrr::transpose(burnedMapList$out)
  do.call(rbind, burnedMapList$LM)
  burnedMapList$burnedMap <- lapply(burnedMapList$burnedMap, terra::unwrap)

  bm <- terra::rast(burnedMapList$burnedMap) |> terra::trim()
  gg <- ggplot() +
    geom_spatraster(data = bm) +
    scale_fill_grass_c("reds") +
    facet_wrap(~lyr) +
    labs(title = sprintf("Fire size: %.0e", fireSize)) +
    theme_void() +
    theme(legend.position = "none")
    
  f_gg <- file.path(figDir, sprintf("%s_fireSize_1e%d.png", format(Sys.Date(), "%Y-%m-%d"), i))
  ggsave(f_gg, gg)
  
  print(gg)
}
```

### Cleaning up

```{r cleanup}
parallel::stopCluster(cl)
gc()

terra::tmpFiles(remove = TRUE)
unlink(cacheDir, recursive = TRUE)
```

### Code and data availability

Code available from <https://github.com/PredictiveEcology/LandMine>.

### Links to other modules

Originally developed as part of the [LandWeb](https://github.com/PredictiveEcology/LandWeb) project.
